### 2011-02-09 paper/learningMethods.tex

There are a plethora of methods available for the creation of the
vector motion generating function $g(t)$ introduced in
\secref{problemDefinition}, and we've chosen two for comparison in
this study.  The first method we evaluated was to use a family of
parametrized gaits, described in this section and the next, and the
second method was to use an evolved non-parametric gait, detailed in
\secref{hyperNeatMethod}.

By a \emph{parametrized gait}, we mean a gait produced by a
parametrized function $g(t; \vec{\theta})$. Fixing the parameter
$\vec{\theta}$ yields a deterministic motion function over time.  We
tried several parametrizations on the robot and, upon obtaining
reasonable early success, settled on one particular parametrization.
We dubbed this the \emph{SineModel5}, as it has five parameters and
employs a sine wave as its root pattern.

The five parameters in $\vec{\theta}$ are as follows:

@@

If we choose a value for the five dimensional parameter
$\vec{\theta}$, then a given physical trial gives us one measurement
of the fitness $f(\vec{\theta})$ of that parameter vector.  Two
things make learning difficult.  First, each evaluation of
$f(\vec{\theta})$ is expensive, taking 15-20 seconds of time on
average.  Second, the fitness returned by such evaluations has proved
to be very noisy, with the standard deviation of the noise often being
roughly equivalent to the size of the measurement.

For both of these reasons, an intelligent \emph{learning algorithm} is
needed.
In this context, a learning algorithm is a method for choosing
%\begin{enumerate}
%\item The value for $\vec{\theta}$ for the initial trial, and
the next value of $\vec{\theta}$ to try, given a list of the
$\vec{\theta}$ values tried so far and the fitness $f(\vec{\theta})$
measured for each.
%\end{enumerate}

We evaluated a total of seven subtly or not so subtly different
learning methods for the parametrized motion models.  All methods
needed a way of picking a $\vec{\theta}$ for the first trial.  In
general this is done by selecting a random starting vector within the
allowable domain.  However, to most directly compare learning methods
by their ability to improve on the initially chosen $\vec{\theta}$, we
decided to evaluate the different methods by starting them at the same
initial $\vec{\theta}$ vectors.

@@

\item \emph{Random}: This method randomly generates parameter vectors
  in the allowable range for every trial.  This strategy was used only
  as baseline for comparison.

>>>>

We have chosen to implement two classes of algorithms to automatically create the motion-generating function $g(t)$ (described in
\secref{problemDefinition}).  The first method uses a family of
parametrized gaits, described in this section and the next. The
second method evolves a non-parametric gait controlled by neural networks, and is detailed in
\secref{hyperNeatMethod}.

By a \emph{parametrized gait}, we mean a gait produced by a
parametrized function $g(t; \vec{\theta})$. Fixing the parameter
$\vec{\theta}$ yields a deterministic motion function over time.  We
tried several parametrizations on the robot and, upon obtaining
reasonable early success, settled on one particular parametrization, which we call  \emph{SineModel5}. Its root pattern is a sine wave and it has five parameters (Table~\ref{tab:params}) .

@@

If we choose a value for the five dimensional parameter
$\vec{\theta}$, then a given physical trial gives us one measurement
of the fitness $f(\vec{\theta})$ of that parameter vector.  Two
things make learning difficult.  First, each evaluation of
$f(\vec{\theta})$ is expensive, taking 15-20 seconds on
average.  Second, the fitness returned by such evaluations has proved
to be very noisy, with the standard deviation of the noise often being
roughly equivalent to the size of the measurement.

We test the ability of different \emph{learning algorithms} to choose the next value of $\vec{\theta}$ to try, given a list of the
$\vec{\theta}$ values already evaluated and their fitness measurements $f(\vec{\theta})$.

We evaluated a total of seven different
learning algorithms for the parametrized motion models.  All methods
needed a way of picking a $\vec{\theta}$ for the first trial of each run.  This could be done by selecting a random starting vector within the
allowable domain.  However, to most directly compare learning methods
by their ability to improve on the initially chosen $\vec{\theta}$, we
decided to evaluate the different methods by starting them at the same three
initial $\vec{\theta}$ vectors in the three runs.

@@

\item \emph{Random}: This method randomly generates parameter vectors

  in the allowable range for every trial.  This strategy was used only
  as baseline for comparison.

### 2011-02-09 paper/results.tex

As mentioned above, the SVM strategy involved many tunable
parameters. The version of SVM tuned for simulation did not show
competitive performance on the real robot. We tried tuning some
parameters on the real robot, but after some amount of tuning, the
method still exhibited too little exploration and easily became stuck
in local minima.

\edit{say something about tuning SVM and the sim figure}

\acmFig{svm_sim_results}{1}{Results for the SVM regression strategy
  in simulation.  This simulation was used to tune the SVM strategy's
  parameters before trying it on the physical robot.  The strategy
  quickly approaches the maximum simulated fitness of 100.}

>>>>

%As mentioned above, the SVM strategy involved many tunable
%parameters. The version of SVM tuned for simulation did not show
%competitive performance on the real robot. We tried tuning some
%parameters on the real robot, but after some amount of tuning, the
%method still exhibited too little exploration and easily became stuck
%in local minima.
%
%\edit{say something about tuning SVM and the sim figure}
%
%\acmFig{svm_sim_results}{1}{Results for the SVM regression strategy
%  in simulation.  This simulation was used to tune the SVM strategy's
%  parameters before trying it on the physical robot.  The strategy
%  quickly approaches the maximum simulated fitness of 100.}
